---
title: "Temperature data"
author: "Paul Rougieux"
date: "21 August 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Sources of historical temperature data for Nancy

Historical temperature data is available :

* on Meteo France's [open data website](https://donneespubliques.meteofrance.fr), in a file called [données SYNOP essentielles OMM](https://donneespubliques.meteofrance.fr/?fond=produit&id_produit=90&id_rubrique=32)
* on [ncdc.noaa.gov](https://www.ncdc.noaa.gov), under the [Essey, FR weather station](https://www.ncdc.noaa.gov/cdo-web/datasets/GHCND/stations/GHCND:FRM00007180/detail)
    * data can be ordered per email, with a maximum of 500 pages per PDF document. 
    * data can be accessed through an API, [with an authentication key/web services token](https://www.ncdc.noaa.gov/cdo-web/token)

Reading temperature data into R:

 * [RNOAA](https://github.com/ropensci/rnoaa) interface to NOAA APIs

# Reading temperature data with the RNOAA package
The package requires a NOAA API key for authentication 
[explanations on how to get and use an API key](https://ropensci.org/blog/2014/03/13/rnoaa/#api-keys---authentication).

Once a NOAA token has been obtained, information can be loaded from the API with curl
([as explained on gis.stackexchange](https://gis.stackexchange.com/a/156657/67897))
for example all datasets available from the Essey station :

    curl -H "token:<enter token here>" https://www.ncdc.noaa.gov/cdo-web/api/v2/datasets?stationid=GHCND:FRM00007180



```{r}
library(rnoaa)
ncdc_stations(stationid='GHCND:FRM00007180')

```

In what unit are precipitation values, are they in Fahrenheit?
I could load historical data from the French website and compare temperatures to find out. 

```{r}
out <- ncdc(datasetid='GHCND', stationid = 'GHCND:FRM00007180', startdate = '2017-05-01', enddate = '2017-05-10')
dtf <- out$data
ncdc_plot(out)
library(ggplot2)
ggplot(dtf , aes(x = date, y = value)) + geom_point() +
    facet_wrap(~datatype)
```

