---
title: "Wine"
author: "Paul Rougieux"
date: "23 January 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(FactoMineR)
library(dplyr)
```

# Load data
```{r}
red <- read.csv("https://archive.ics.uci.edu/ml/machine-learning-databases/wine-quality/winequality-red.csv", sep=";")
white <- read.csv("https://archive.ics.uci.edu/ml/machine-learning-databases/wine-quality/winequality-white.csv", sep=";")
```

# Explore
## Acidity in red and white
White wines have a lower pH than red wines.
```{r}
summary(red$pH)
ggplot(red, aes(x = pH)) +
    geom_histogram() +
    ggtitle("Red wine")

summary(white$pH)
ggplot(white, aes(x = pH)) +
    geom_histogram() +
    ggtitle("White wine")
```


## Alcohol and quality
It seems that alcohol is positively correlated with quality.
```{r}
# plot structure with NULL data, data will be specified later in the geom_point call.
p <- ggplot(NULL, aes(x = as.factor(quality), y = alcohol))
    
p + geom_boxplot(data = red) + ggtitle("Red wine")
p + geom_boxplot(data = white) + ggtitle("White wine")
```


## Sulphates, sulfur.dioxide and quality
Total.sulfur.dioxide is negatively correlated with quality.
```{r}
# plot structure with NULL data, data will be specified later in the geom_point call.
p <- ggplot(NULL, aes(x = sulphates, y = total.sulfur.dioxide, color = as.factor(quality))) +
    facet_grid(quality ~ .)
    
p + geom_point(data = red) + ggtitle("Red wine")
p + geom_point(data = white) + ggtitle("White wine")
```




# ACP
sthda.com: [Méthodes des Composantes Principales dans R: Guide Pratique](http://www.sthda.com/french/articles/38-methodes-des-composantes-principales-dans-r-guide-pratique/73-acp-analyse-en-composantes-principales-avec-r-l-essentiel/)

## Red wine
```{r}
library(FactoMineR)
res.pca <- PCA(select(red, -quality))
res.pca <- PCA(red)
summary(res.pca)
```

## White wine
```{r}
library(FactoMineR)
res.pca <- PCA(white)
summary(res.pca)
```



# Correlation
```{r}
cor(red)
cor(white)
```

